---
title: "ENV S 193DS Visualization"
author: "Arantza Lira"
format: html
toc: true
execute: 
  message: false #messages dont show up
  warning: false #warnings dont show up
---

During ENV S 193 DS, we had three homework assignments using data from ecological research and pollution data. This visualization is from homework 2. It uses data from the 2017-2018 Thomas Fire that burned 281,893 acres in Santa Barbara County and Ventura County areas. The figure below demonstrates the concentration of particulate matter in the Goleta and Santa Barbara area after the fire.

### Data set up

```{r}
# installing and reading in packagaes
library("tidyverse")
library("janitor")
library("plotly")

# Reading in Thomas fire data
sbpm <- read_csv("C:/Users/arant/Desktop/ENV S 193DS-R/arara4191.github.io/posts/stats/sbpm.csv")

#data wrangling to seperate data points by location
gol_sb <- sbpm %>%
filter(local_site_name %in%
c("Goleta", "Santa Barbara"))


```

### Visualization Code

```{r}
#creating markers
data_marker <- gol_sb |>  # calling dataframe 
  # create a new column/field called marker 
  mutate(marker = paste0("date: ", date, "<br>",# quoted words show the type of infor that will be displayed : data
                         # <br> creates a line break in marker dispaly
                         "PM Concentration: ", pm2_5))



thomas_plot <- ggplot(data = data_marker, # calling dataframe
aes(x = local_site_name, # x-axis data
y = pm2_5, # y-axis data
color = local_site_name, # coloring by site
shape = local_site_name)) +
geom_boxplot(aes(text = NULL)) + # first layer should be a boxplot
geom_jitter(aes(text = marker),width = 0.2, height = 0) + # making sure points don't jitter vertically
labs(x = "Location", # labelling the x-axis
y = "2.5 Concentration (ug/m^3 LC)",
color = "Location Key") + # labeling the y-axis
# changing appearance: colors, labels, and theme
scale_color_manual(values = c("Santa Barbara" = "purple",
"Goleta" = "orange")) +
theme_get()+
theme(legend.position = "none") # adding theme and legend condition


#plotly package to make figure interactive
# editing display
thomas_int <- ggplotly(thomas_plot, tooltip = "text") %>%
  plotly::layout(
    font = list(family = "Times"),# editing the font for the marker hover-tex
    hoverlabel = list(
      hovermode = "closest",
      font = list(
        family = "Times",
        size = 13,
        color = "#FFFFFF"# font color
      ),
      align = "left" #indent
    )
  )

```

### Thomas Fire Visualization

```{r}
#displaying plot
thomas_int
```
